---
layout: post
title: "A Custom `ggplot2` Theme for [Facebook](https://www.facebook.com/)"
date: "`r Sys.time()`"
categories: blog, fun
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Why? 

I thought it would be fun to create a Facebook theme for `ggplot2` after seeing [this (kind of old) article](http://www.theatlantic.com/technology/archive/2014/02/when-you-fall-in-love-this-is-what-facebook-sees/283865/?utm_source=mitfb) where Facebook data analysts have clearly used [`ggplot2`](http://docs.ggplot2.org/current/) to make their graphs. I [tweeted about it]() and because I'm "on winter break" right now and I just revamped my website, I decided to create this blog post about it.

I use the [`ggplot2` themes vignette](http://docs.ggplot2.org/dev/vignettes/themes.html) and the package [`ggthemes`](https://github.com/jrnold/ggthemes) as references when creating this Facebook theme.

### The Theme

```{r fbtheme}
library(ggplot2)

# facebook colors
# from https://designpieces.com/2012/12/facebook-colour-palette/
# and darkblue from http://www.color-hex.com/color/3b5998
fb_colors <- c(blue = "#3b5998", medblue = "#6d84b4", 
               lightblue = "#afbdd4", lightestblue = "#d8dfea", 
               white = "#ffffff", darkblue = "#111a2d")

# facebook color palette (see ggthemes)
fb_pal <- function () {
  function(n) {
    colors <- fb_colors[rev(c("darkblue", "blue", "lightblue"))]
    unname(colors[seq_len(n)])
  }
} 

# discrete scale colors for facebook
scale_colour_fb <- function (...) {
    discrete_scale("colour", "fb", fb_pal(), 
                   ...)
  }  
scale_color_fb <- scale_colour_fb


theme_fb <- function(base_size = 11, base_family = "Lucida Grande", 
                     fbcolors = fb_colors){
  half_line <- base_size/2
  theme_bw() %+replace% 
    theme(text = element_text(family = base_family, face = "plain", 
                              color = fbcolors["medblue"], size = base_size,
                              hjust = .5, vjust = .5, angle = 0, lineheight = 1.1, 
                              margin = margin(), debug = FALSE ),
          panel.grid.major = element_line(colour = fb_colors["lightestblue"]),
          panel.grid.minor = element_line(colour = fb_colors["lightestblue"],
                                          size = .25), 
          plot.title = element_text(size = rel(1.2), hjust = 0, 
                                   vjust = 1, 
                                   margin = margin(b = half_line * 1.2), 
                                   face = "bold", color = fb_colors["blue"]),
          panel.border = element_rect(fill = NA, 
                                      colour = fb_colors["lightblue"]),
          strip.background = element_rect(fill = fb_colors["medblue"], 
                                          colour = fb_colors["lightblue"]),
          strip.text = element_text(colour = fb_colors["white"], 
                                    size = rel(0.8)),
          axis.text = element_text(size = rel(0.8), 
                                   colour = fb_colors["darkblue"]),
          axis.ticks = element_line(colour = fb_colors["darkblue"])
  )
}
```

### Using the Facebook Theme

```{r fbex, fig.align='center', fig.height=5}
data("mtcars")

ggplot(data = mtcars, aes(x = hp, y = mpg)) + 
  geom_point(aes(color = as.factor(cyl))) + 
  facet_wrap(~gear, labeller = label_both) + 
  labs(title = "Cars Data, Facebook Theme") + 
  scale_color_fb() + 
  theme_fb()
```

