\documentclass[margin,line]{res}
\usepackage{multicol}
\usepackage[bottom=.65in,top=.65in]{geometry}
\usepackage{amsmath}
\usepackage{url}
\usepackage{enumitem}
\usepackage{etaremune}

\oddsidemargin -.5in
\evensidemargin -.5in
\textwidth=6.0in
\itemsep=0in
\parsep=0in

\newenvironment{list1}{
  \begin{list}{\ding{113}}{%
      \setlength{\itemsep}{0in}
      \setlength{\parsep}{0in} \setlength{\parskip}{0in}
      \setlength{\topsep}{0in} \setlength{\partopsep}{0in} 
      \setlength{\leftmargin}{0.17in}}}{\end{list}}
\newenvironment{list2}{
  \begin{list}{$\bullet$}{%
      \setlength{\itemsep}{0in}
      \setlength{\parsep}{0in} \setlength{\parskip}{0in}
      \setlength{\topsep}{0in} \setlength{\partopsep}{0in} 
      \setlength{\leftmargin}{0.2in}}}{\end{list}}


\begin{document}

\name{Samantha C. Tyner, PhD \vspace*{.1in}}

\begin{resume}
\section{\sc Contact}
\vspace{.05in}
{{\it Phone:}  630.346.3539 $\quad$ {\it E-mail:} sctyner90@gmail.com $\quad$ {\it Website:} \urlstyle{same}\url{sctyner.github.io} }
%\begin{tabular}{@{}p{2in}p{4in}}
%207 Todd Circle  & {\it Phone:}  630.346.3539 \\            
%Ames, Iowa USA 50011  &  {\it E-mail:} sctyner@iastate.edu\\      
%& {\it Website:} \urlstyle{same} \url{sctyner.github.io}
%\end{tabular}

\section{\sc Education}
{\bf Iowa State University}, Ames, Iowa, USA\\
\vspace*{-.1in}
\begin{list1}
\item[] {\bf Postdocoral Research Associate, CSAFE} \hfill {\bf December 2017 - Present}
\item[] $\quad$ Principal Investigator: Alicia Carriquiry
\item[] {\bf Ph.D., Statistics}  \hfill {\bf December 2017}
\item[] $\quad$ Advisor: Heike Hofmann
\item[] $\quad$ Dissertation: \textit{Graphical discovery in stochastic actor-oriented models \\  \phantom{AAAAAAAA} for social network analysis}
\item[] {\bf M.S., Statistics}  \hfill {\bf May 2015} 
% \item[]Cumulative GPA: 3.63/4.00
\end{list1}
\vspace{-.1in}

{\bf Augustana College}, Rock Island, Illinois, USA\\
%{\em Department of Mathematics} 
\vspace*{-.1in}
\begin{list1}
\item[] {\bf B.A. Mathematics, Economics, and French}, \textit{summa cum laude} \hfill {\bf May 2012}
%\item[]Cumulative GPA: 3.93/4.00
\end{list1}

%\vspace*{-2.5mm}
%NSF Vertical Integration of Research and Education in Statistics and
%Mathematical Sciences\\ (VIGRE) teaching fellowship.
%
%\vspace*{-2.5mm}
%Carleton College: graduated Magna Cum Laude, Honors in Biology, Phi Beta Kappa, 1993

\section{\sc Computing \\ Skills} 

\textbf{Expertise:} R/RStudio, Shiny, Markdown,  Git/Github, HTML/CSS,  \LaTeX \\
\textbf{Familiarity:} Python, Julia, C/C++, SAS, JAGS/WinBugs, Maple, Mathematica

\vspace{-.2cm}


\section{\sc Professional \\ Experience}
{\em Center for Statistics \& Applications in Forensic Evidence (CSAFE)}, \\
{\em Iowa State University}, Ames, Iowa USA

\vspace{-.3cm}

{\bf Postdoctoral Research Associate } \hfill {\bf December 2017 - Present}\\
\vspace{-.3cm}
\begin{itemize}
\setlength\itemsep{0em}
\item Author educational materials for legal professionals on the use of statistics in forensic science
\item Conduct research relating to human factors, statistical computing in forensic science
\item Mentor graduate students on their research projects
\item Coordinate the summer research experience for undergraduates (REU) from partner schools
\end{itemize} 
\vspace{-.3cm}

{\bf Research Assistant} \hfill {\bf August 2015 - December 2017}\\
\vspace{-.3cm}
\begin{itemize}
\setlength\itemsep{0em}
\item Author educational materials for legal professionals on the use of statistics in forensic science
\item Conduct research relating to human factors, statistical computing in forensic science
\end{itemize} 

\vspace{-.3cm}

{\bf REU Graduate Student Advisor} \hfill {\bf May 2017 - August 2017}\\
\vspace{-.3cm}
\begin{itemize}
\setlength\itemsep{0em}
\item Instructed a group of ten students on R coding, statistics, and Github project management 
\item Supervised research projects on automated shoeprint matching and steganalysis
\end{itemize} 

{\em Social and Decision Analytics Laboratory, Biocomplexity Institute, \\ Virginia Polytechnic Institute and State University}, Arlington, VA, USA

\vspace{-.3cm}

{\bf Graduate Student Fellow} \hfill {\bf May 2016 - August 2016}\\
\vspace{-.3cm}
\begin{itemize}
\setlength\itemsep{0em}
\item Participated in the interdisciplinary Data Science for Public Good summer program
\item Led groups of undergraduate researchers on data-driven policy projects
\item Analyzed and combined data from a variety of administrative, census, and GIS sources 
\item Consulted with county staff on projects to improve government services in Arlington, VA.
\end{itemize}


{\em NORC at the University of Chicago}, Chicago, IL, USA

\vspace{-.3cm}

{\bf Graduate Research Assistant} \hfill {\bf August 2015 - April 2016}\\
\vspace{-.3cm}
\begin{itemize}
\setlength\itemsep{0em}
\item Developed a dynamic reporting system to adapt to changes in Illinois' healthcare marketplace  
\item Continued my summer internship work part time as a consultant.  
\end{itemize}

{\bf Data Visualization Intern} \hfill {\bf May 2015 - August 2015}\\
\vspace{-.3cm}
\begin{itemize}
\setlength\itemsep{0em}
\item Created data visualizations for a variety of projects in the Statistics \& Methodology group
\item Designed a multi-use dashboard of education data in Washington state
\item Developed a web app with public health, GIS data to uncover potential epidemic regions
\item Led four hour-long internal seminars on the theory and best practices of data visualization.  
\end{itemize}

\vspace{-.3cm}


{\em Department of Statistics, Iowa State University}, Ames, Iowa USA

\vspace{-.3cm}

{\bf Teaching Assistant} \hfill {\bf August 2014 - May 2015}\\
\vspace{-.3cm}
\begin{itemize}
\setlength\itemsep{0em}
\item Instructor for STAT 104: Introduction to Statistics
\item Taught weekly lectures and labs
\item Wrote and graded exams.
\end{itemize}
 
%{\em Graduate Student} \hfill {\bf August 2012 - Present}\\
%Masters and PhD level coursework. 

{\em Kingland Systems Corporation}, Ames, Iowa USA

\vspace{-.3cm}

{\bf Statistical Analyst Intern} \hfill {\bf May 2014 - August 2014}\\
\vspace{-.3cm}
\begin{itemize}
\setlength\itemsep{0em}
\item Served as a member of the Process Management team 
\item Performed data analysis on project and product data to refine internal processes and tools
\item Built predictive models used in planning and estimating time to complete sofware development
\item Produced graphs and reports to support internal visibility of projects, products, and portfolios. 
\end{itemize}

\vspace{-.1cm}

{\em Institute for Transportation, Iowa State University}, Ames, Iowa USA

\vspace{-.3cm}

{\bf Research Assistant} \hfill {\bf August 2013 - August 2014}\\
\vspace{-.3cm}
\begin{itemize}
\setlength\itemsep{0em}
\item Modeled curve navigation on rural highways using large USDOT data sets. 
\end{itemize}

\vspace{-.1cm}

{\em Department of Statistics, Iowa State University}, Ames, Iowa USA

\vspace{-.3cm}


{\bf Teaching Assistant} \hfill {\bf August 2012 - July 2013}\\
\vspace{-.3cm}
\begin{itemize}
\setlength\itemsep{0em}
\item Graded weekly homework assignments for STAT 104: Introduction to Statistics
\item Supervised lab sections twice weekly.
\end{itemize}


%\vspace{-.1cm}
%
%{\em Valparaiso Experience in Research by Undergraduate Mathematicians (VERUM) \\ Valparaiso University}, Valparaiso, Indiana USA \hfill {\bf May 2011 - July 2011}\\
%\vspace{-.3cm}
%\begin{itemize}
%\setlength\itemsep{0em}
%\item Participated in the VERUM summer program. 
%\item Worked with Dr. Lara Pudwell on pattern avoidance in full binary tree graphs.  
%\end{itemize} 


%\clearpage


%{\bf NORC at the University of Chicago}, Chicago, IL, USA

%\vspace{-.3cm}

\section{\sc Publications}

\begin{etaremune}[start=5, itemindent = -1.25em]

\item \textbf{Tyner, S.}, Briatte, F.,  and Hofmann, H. ``Network Visualization with ggplot2". {\em The R Journal}, 9(1):27-59, June 2017.

\item \textbf{Tyner, S.} ``Using the R Package geomnet: Visualizing Trans-Atlantic Slave Trade of Africans, 1514-1866." \textit{CHANCE}.  Sep 2016. 29(3)

\item \textbf{Tyner, S.}, "The Joy of Clustering." \textit{Significance Magazine Online}. 11 Nov 2015 \urlstyle{same}\url{http://www.statslife.org.uk/culture/2553-the-joy-of-clustering}

\item Hallmark,S., \textbf{Tyner, S.},  Oneyear, N., Carney, C., and McGehee, D. ``Evaluation of driving behavior on rural 2-lane curves using the SHRP 2 naturalistic driving study data." \textit{Journal of Safety Research}, 54 (2015), p17-27.

\item Dairyko, M., Pudwell, L., \textbf{Tyner, S.}, and Wynn, C.  ``Non-Contiguous Pattern Avoidance in Binary Trees" \textit{Electronic Journal of Combinatorics} 19 (3) (2012), P22.

\end{etaremune}


\section{\sc Workshops Taught}

{\bf The Life-Changing Magic of Tidying Data}, Ames, Iowa, USA \hfill {\bf December 8, 2017}\\
Workshop on the principles and process of cleaning and tidying messy data sets with the R packages dplyr, tidyr, and others. Presented at Iowa State University, invited by the R-Ladies Ames group.


{\bf The Week of R}, Ames, Iowa, USA

\vspace{-.4cm}

{\em with Joe Papio, Haley Jeppson, and Heike Hofmann} \hfill {\bf June 12-15, 2017}\\
Intensive workshop on the basics of R, as well as data cleaning, webscraping, data visualization, and Shiny in R and RStudio. Presented at Iowa State University.

{\bf Webscraping with tidyverse Packages}, Ames, Iowa, USA \hfill {\bf February 9, 2017}\\
Workshop on the process of scraping data from the web using the R packages rvest, xml, and others. Presented at Iowa State University, invited by the R-Ladies Ames group.

{\bf Empowering Women with Self-defense Skills using R}, Charlotte, NC, USA

\vspace{-.4cm}

{\em with Heike Hofmann} \hfill {\bf October 22, 2016}\\
Workshop on data cleaning, webscraping, and data visualization in R. Presented at the 2016 Women in Statistics and Data Science conference.

\vspace{-.2cm}

\clearpage 
{\bf The Week of R}, Ames, Iowa, USA

\vspace{-.4cm}

{\em with Eric Hare, Andee Kaplan, and Heike Hofmann} \hfill {\bf May 16-19, 2016}\\
Intensive workshop on the basics of R, as well as data cleaning, webscraping, data visualization, and Shiny in R and RStudio. Presented at Iowa State University.
 

%\section{\sc Conference Presentations} 
%
%{\bf Joint Statistical Meetings}, Chicago, Illinois, USA
%
%\vspace{-.3cm}
%{\em Topic-Contributed Session} \hfill {\bf August 2016}\\
%``Visual Inference for Networks" with Hofmann, H..
%
%
%{\bf Annual Conference of the Royal Statistical Society}, Exeter, Devon, UK
%
%\vspace{-.3cm}
%{\em Invited Talk - Young Statisticians Section} \hfill {\bf September 2015}\\
%``The Joy of Clustering (with Bob Ross)" 
%
%
%{\bf Joint Statistical Meetings}, Seattle, Washington, USA
%
%\vspace{-.3cm}
%{\em Contributed Session} \hfill {\bf August 2015}\\
%``Network Visualization in \texttt{ggplot2}: \texttt{geom\_net}" with Hofmann, H..
%
%{\bf Women in Statistics}, Cary, North Carolina, USA
%
%\vspace{-.3cm}
%{\em Poster Session} \hfill {\bf May 2014}\\
%``Using Change Point Models to Assess Drivers' Navigation of Rural Curves" with Hallmark, S.
%
%{\bf Joint Statistical Meetings}, Montr\'{e}al, Qu\'{e}bec, CAN
%
%\vspace{-.3cm}
%{\em Data Expo Poster Session} \hfill {\bf August 2013}\\
%``An Illustration of the Soul of the Community" with Grant, M.  
%
%{\bf Joint Mathematics Meetings}, Boston, Massachusetts, USA
%
%\vspace{-.3cm}
%{\em Undergraduate Poster Session} \hfill {\bf January 2012}\\
%``Non-Contiguous Pattern Avoidance in Binary Trees" with Dairyko, M. and Wynn, C. 
%
%{\bf Illinois Section of Mathematical Association of America}, Naperville, Illinois, USA
%
%\vspace{-.3cm}
%{\em Undergraduate Presentation Session} \hfill {\bf March 2011}\\
%``Logarithmic Regression and the Phillips Curve" 
%

%\section{\sc Computer \\Skills} 
%\textbf{Mathematical Computing}:  R, RStudio, SAS, Python, Julia, Mathematica, Maple, WinBugs \\
%\textbf{Other}: \LaTeX, \texttt{knitr}, Markdown, Microsoft Office

\section{\sc Leadership / Service}

{\em Iowa State University}, Ames, Iowa USA

\vspace{-.3cm}

{\bf Co-founder and co-organizer, R-Ladies Ames} \hfill {\bf December 2016 - Present}\\
Founded the Ames chapter of the global organization R-Ladies, which promotes gender parity among R users, R package authors, and members of the R foundation and board of directors. 

\vspace{-.3cm}

{\bf Departmental Committees} \hfill {\bf August 2014 - May 2017}\\
Served alongside faculty and staff as the student member of departmental committees: the Social Committee (2014-15) and the Honors and Awards Committee (2015-16, 2016-17). 

\vspace{-.3cm}

{\bf The Iowa STAT-ers Club} \hfill {\bf August 2012 - May 2016}\\
Led this Statistics Department graduate student organization in multiple executive positions: Secretary (2013-14), Social Chair (2014-15), and President (2015-16).

\section{\sc Honors \& \\Awards} 
\begin{description}[itemindent = -2.75em]
\setlength\itemsep{-1em}
\item \textit{Winner, rOpenSci 2018 Fellowship} for proposal ``Open Forensic Science in R" \urlstyle{same}\url{ropensci.org}\\
\item \textit{Winner, Statistical Computing and Statistical Graphics Sections of the ASA 2016 Student Paper Competition} for ``Using the geomnet Package: Visualizing African Slave Trade, 1514 - 
1866"\\
\item \textit{Runner up, 2015 Young Statisticians Writing Competition, by the Young Statisticians Section of the Royal Statistical Society} for ``The Joy of Clustering (with Bob Ross)"\\
\item \textit{Alumni Fellowship}, Dept. of Statistics, Iowa State University, Awarded Fall 2012\\
\item \textit{Phi Beta Kappa}, Inducted May 2012\\
\item \textit{Pi Mu Epsilon}, Inducted May 2011\\
\item \textit{Omicron Delta Kappa}, Inducted December 2011\\
\item \textit{Augustana College Dean's List}, all terms Fall 2008 - Spring 2012\\
\end{description}

\vspace{-.3cm}





\end{resume}
\end{document}




